What's in a name? On the (automated) topic naming of software maintenance activities

### 2010-01-27 icpc/icpc2010.tex
updated methodology somewhat

In this paper we use LDA to identify topics. The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate. The algorithm returns a word-document matrix relating the occurrence frequency of a

@@

\section{Methodology}
We leveraged the data we gathered in \cite{Hindle09ICSM} for this work. To gather the data, we first extract the commit log comments from a
project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
into 30-day non-overlapping windows. Each window is subject to LDA and then we analyze and visualize the results. %Figure \ref{fig:commits}
depicts the general process for processing and analyzing the commit messages.

Find two projects with open, accessible source code in similar fields. We selected MySQL and MaxDB, open-source, near-commercial quality database software. We used source code commit data (the message attached to the commit). A typical message might look like ``". From that dataset, we created an XML file which grouped commits into temporal periods, each approximately X weeks wide. Given a text file full of commits, indexed by date and time, we then ran Latent Dirichlet Allocation ~\cite{Blei2003} on the data. LDA works by creating topic models -- word frequency distributions -- for each period. A topic model is a statistical summary of the important subjects for that dataset. E.g., in Period 1, the most important topic might have been ``security", with ``sql'' second, down on to less relevant topics. The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists. These approaches are unsupervised in that they rely solely on the source data with no human intervention.

Once we have a list of topics for each period, we generated some word lists to assess to what extent our topics fit word lists. Intuitively, we would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period. We use several different word lists. One we tried was 9126, the ISO quality model. If words in the quality model were also found in a period, we automatically assigned that period to that subject (e.g., security, performance, reliability).

@@

\subsection{Project characteristics}
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.5, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's `SLOCCount'.}. There are approximately ... developers working on it. MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code. The MySQL project started in 1994.

>>>>

In this paper we use LDA to identify topics.  LDA works by creating topic models for each period. A topic model is a statistical summary of the important subjects for that dataset.  The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate. The algorithm returns a word-document matrix relating the occurrence frequency of a 
 The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists. These approaches are unsupervised in that they rely solely on the source data with no human intervention.

@@

\section{Methodology}
 % To gather the data, we first extract the commit log comments from a
% project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
% into 30-day non-overlapping windows. Each window is subject to LDA analysis. %Figure \ref{fig:commits}
% depicts the general process for processing and analyzing the commit messages.

We wanted to compare two projects with open, accessible source code in similar functional areas. This allows us to partially control for differences in functional requirements. We selected MySQL and MaxDB, open-source, near-commercial quality database software. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.5, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's `SLOCCount'.}. There are approximately ... developers working on it. MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code. The MySQL project started in 1994.

From each project, we used source code commit data (the message attached to the commit). We leveraged the data we gathered in \cite{Hindle09ICSM} for this work. A typical commit message is: ``history annotate diffs bug fixed (if mysql_real_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course) ". We extracted these messages and indexed them by creation time.

From that dataset, we created an XML file which grouped commits into temporal periods, each approximately 4 weeks wide. Given a text file full of commits, indexed by date and time, we then ran an LDA modeler on each period, to generate a list of topics per period. E.g., in Period 1, the most important topic might have been ``security", with ``sql'' second, down on to less relevant topics. We arbitrarily set the number of topics to generate to 20, based on past experiences.

	Once we have a list of topics for each period, we compared them with lists of keywords to assess to what extent our topics fit word lists. Intuitively, we would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period. We used several different word lists. One we tried was 9126, the ISO quality model~\cite{iso9126}. It describes six high-level quality requirements (listed in Table \ref{}). There is some debate about the significance and importance of the terms in this model. However, it is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' which is sufficient for the purposes of this research. %exp3

	We then extended our word lists with related words using Wordnet.

	 If words in the quality model were also found in a period, we automatically assigned that period to that concept (e.g., security, performance, reliability).

@@
